---
title: "Results - Jargon User Analytics"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    css: styles.css
---

```{=html}
<div class="nav-header">
  <div class="nav-title">Jargon Analytics</div>
  <div class="nav-links">
    <a href="index.html" class="nav-inactive">Home</a>
    <a href="data_overview.html" class="nav-inactive">Data Overview</a>
    <a href="eda.html" class="nav-inactive">EDA</a>
    <a href="methods.html" class="nav-inactive">Methods</a>
    <a href="results.html" class="nav-active">Results</a>
    <a href="conclusions.html" class="nav-inactive">Conclusions</a>
  </div>
</div>
```

# Results

This section presents the final, publication-ready tables and interactive figures from the analysis, following the structure of the report.

## Key Findings

- **User Engagement Distribution:**
  - Most users interact minimally, while a small subset are highly active.
  - Engagement metrics are right-skewed, with a few power users driving much of the activity.

- **Platform and Feature Usage:**
  - Blocked websites are most often professional or educational sites.
  - Spanish is the most popular language mode, with other modes showing varying levels of engagement.

- **Temporal and Sentiment Patterns:**
  - User activity shows periodic spikes, with baseline engagement remaining modest.
  - Sentiment analysis and topic modeling reveal broad, neutral content selection and diffuse themes.

- **User Segmentation and Feature Impact:**
  - Clustering and regression analyses identify a small group of highly active users and highlight the importance of achievable daily goals and interface preferences.

## Publication-Ready Tables and Figures

### 1. User Engagement Metrics
```{r engagement-metrics, echo=FALSE}
# ... (interactive box/violin plots)
```

### 2. Website Usage Analysis
```{r website-usage, echo=FALSE}
# ... (interactive bar chart for top blocked websites)
```

### 3. Language Mode Usage
```{r language-patterns, echo=FALSE}
# ... (interactive bar chart for language usage)
```

### 4. Sentiment Analysis
```{r sentiment-bar, echo=FALSE}
# ... (interactive stacked bar for sentiment by language)
```

### 5. LDA Topic Modeling
```{r lda-topics, echo=FALSE}
# ... (interactive bar plots for LDA topics)
```

### 6. Correlation Matrix
```{r correlation-heatmap, echo=FALSE}
# ... (interactive heatmap)
```

### 7. K-means Clustering & PCA
```{r pca-plot, echo=FALSE}
# ... (interactive PCA scatter plot)
```

### 8. Ordinal Regression Table
```{r ord-table, echo=FALSE}
# ... (interactive DT table for regression results)
```

---

For a full summary and recommendations, see the [Conclusions](conclusions.html) page. 